# Programming {#programming}

## R

```{r setup, echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(patchwork)
library(palmerpenguins)

# default theme for ggplot
theme_set(theme_bw())

# do not output errors and warnings in Markdown
knitr::opts_chunk$set(error = FALSE, warning = FALSE)
```


### Graphics

#### ggplot2

##### Defaults

```{r, eval=FALSE}
# Default theme
theme_set(theme_bw())
```


##### Save to file system

```{r, eval=FALSE}
plt_spec_depth <- 
  penguins %>% 
  ggplot(aes(x = species, y = bill_depth_mm)) +
  geom_boxplot()

ggsave(here("figs", "plt_spec_depth.tiff"), plot = plt_spec_depth, device = "tiff", 
       width = 5, height = 5, dpi = 300, compression = "lzw")

plt_spec_depth
```

##### Rotate axis lables

```{r}
penguins %>% 
  ggplot(aes(x = species, y = bill_depth_mm)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

##### Zoom axis limits

```{r}
no_zoom <- 
  penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()+
  ggtitle("No zoom")

zoom <- 
  penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  coord_cartesian(xlim = c(40,50)) +
  ggtitle("Zoom x-axis to c(40,50)")

no_zoom + zoom
```

### Reporting

#### R Markdown

##### Defaults

```{r, eval=FALSE}
# Do not output code
knitr::opts_chunk$set(echo = FALSE)

# Do not output errors and warnings
knitr::opts_chunk$set(error = FALSE, warning = FALSE)

# Produce self-contained html files
knitr::opts_chunk$set(cache.path = 'report_cache/')
```

I didn't find the source for the self-contained html file anymore, but this helped producing one at least once. 
Should be more like a last resort than a default, though.

##### File encoding

Blog post by Yihui Xie, why every R Markdown file should be encoded in UTF-8: [UTF-8, and UTF-8 only, or we cannot be friends](https://yihui.org/en/2018/11/biggest-regret-knitr/)

### Objects

#### Description

##### Number of unique values per variable in a data frame

This is like `table()`, and loops over each variable of a data frame.
Using *tidy evaluation*, code from the dplyr vignette [Programming with dplyr](https://dplyr.tidyverse.org/articles/programming.html). 


```{r, eval=FALSE}
for (var in names(penguins)) {
  penguins %>% count(.data[[var]]) %>% print()
}
```


#### Compare objects

##### Hash functions

To compare whether to files on the file system are identical, the MD5 hash function can be used (other hash functions as SHA1, SHA256 etc. could be used as well, bud MD5 should be one of the faster ones).
To not just hash the path, a connection to the file has to be established (with `file()`).
`file()` has several `modes` in which reading and writing can be done.
The hashed values will be different depending which mode is used.

> Whether the hashed values of the different files match may depend on the used mode.

Source: [Link](https://stackoverflow.com/questions/59833931/md5-file-hash-functions-in-r-returning-different-values)

```{r, eval=FALSE}
md5_orig <- md5(file(orig, open = "rt"))
md5_rebuild <- md5(file(rebuild, open = "rt"))
  
identical(md5_orig, md5_rebuild)
```

###### function which takes two file paths as input

```{r}
comp_files_md5 <- function(orig, rebuild) {
  con_orig <- file(orig, open = "rt")
  md5_orig <- md5(con_orig)
  con_rebuild <- file(rebuild, open = "rt")
  md5_rebuild <- md5(con_rebuild)
  
  close(con_orig)
  close(con_rebuild)
  
  identical(md5_orig, md5_rebuild)
}
```


##### Visual comparison

R objects can be visually compared with the package `diffobj`.

```{r, eval=FALSE}
mod <- penguins
mod$sex <- sample(mod$sex, length(mod$sex))
diffobj::diffObj(penguins, mod)
```

## Regular expression

### Find 'word' not followed by a certain character

> keywords: #regex

Source: [Link](https://stackoverflow.com/questions/31201690/find-word-not-followed-by-a-certain-character)

The (?!@) negative look-ahead will make word match only if @ does not appear immediately after word:

```{}
word(?!@)
```

